% !TEX root = ../thesis-example.tex
%
\pdfbookmark[0]{Acknowledgement}{Acknowledgement}
\chapter*{Acknowledgement}
\label{sec:acknowledgement}
\vspace*{-10mm}


These three years have been a long journey with many good times and also some hard moments, which I
lived together with many people that I wish to thank.

First of all I would like to thank Massimo Masera and Stefania Bufalino, not only for their 
supervision but also for the extraordinary support and the opportunity they gave me to work in the
ALICE group of Torino.
I would like especially to thank Maximiliano Puccio, who followed and supported my work starting from
my CERN experience in 2016, solving my doubts and discussing with me till 9.00 p.m. about my thesis.

I should also like to say a big thank-you to all the people of the ALICE group of Torino and
in particular the PhD students: Stefano Trogolo, Luca Barioglio, Matteo Concas, Fabrizio Grosa and 
Ivan Ravasenga, it has been a pleasure to share with you the office and the coffee breaks!

Thanks are due also to all my friends of the Physics Department, in particular Eugenia Cammalleri,
Luca Candela, Debora Giolo and Francesco Schepis for sharing with me joy and sorrow of these years
of study.

I would also like to say a special word of thanks to Dr. Barbara Fornas, without whose help
I wouldn't be here today, finishing to write this thesis.

Finally I would like to thank my parents for giving me the possibility to study in Torino
always unreservedly supporting me.
And, of course, I have to thanks my sisters Anna and Elena for putting up with me in these years.

\textit{Torino, 15 October 2018} \begin{flushright} P. F. \end{flushright}